\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-19 tex/paper.tex
save

Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
Let consider the multivariate random variables $X_z: \Omega \mapsto
\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
on a nuisance parameter $Z$ whose values $z \in {\cal Z}$  define a
parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
induce together a joint probability distribution $p(X,Y|z)$, where the
conditional on $z$ denotes $X_z$. For training, let further assume a finite set $\{
x_i, y_i, z_i \}_{i=1}^N$ of realizations $X_{z_i}(\omega_i), Y(\omega_i)$, for
$\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
is to learn a function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal Y}$ of
parameters $\theta_f$ (e.g., a neural network-based classifier if ${\cal Y}$ is
a finite set of classes) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
the cross-entropy). In addition, we require that $f(X_z ; \theta_f)$ should be
robust to the value $z$ of the nuisance parameter  -- which remains unknown at
test time. More specifically, we aim at building $f$ such that in the ideal case
\begin{equation}\label{eqn:criterion-true}
f(X_{z}(\omega) ; \theta_f) = f(X_{z^\prime}(\omega) ; \theta_f)
\end{equation} for all
samples $\omega \in \Omega$ and all $z, z^\prime$ pairs of values of the
nuisance parameter.

@@

    \node[above] at (12.75,-2.5) {$\mu_{\theta_r}(f(X;\theta_f))$};
    \path[-o, shorten >= 0pt] (11,-2.5) edge (14,-2.5);
    \node[above] at (12.75,-3.5) {$\sigma_{\theta_r}(f(X;\theta_f))$};
    \path[-o, shorten >= 0pt] (11,-3.5) edge (14,-3.5);

@@

    \node[neuron] (r-O1) at (7cm+4*\layersep,-2.5cm) {};
    \node[neuron] (r-O2) at (7cm+4*\layersep,-3.5cm) {};

@@

    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O1);
    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O2);

@@

    \draw (14,-1.5) rectangle (17,-4.5);
    \path[->, shorten >= 0pt] (15.5,-0.5) edge (15.5,-1.5);
    \node[above] at (15.5,-0.5) {$Z$};
    \path[->, shorten >= 0pt] (15.5,-4.5) edge (15.5,-5.5);
    \node[below] at (15.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
    \node at (15.5,-3) {Truncated ${\cal N}$};

@@

    \caption{Architecture for the adversarial training of a binary classifier
    $f$ against a continuous parameter $Z$, assuming that $Z|f(X;\theta_f)$
    follows a truncated gaussian distribution.}
    \label{fig:architecture}
    \end{figure*}

>>>>

Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
Let consider the multivariate random variables $X_z: \Omega \mapsto
\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_z$ denotes a dependence
of the functional $X$ on a nuisance parameter $Z$ whose values $z \in {\cal Z}$  define a
parameterized family of its systematic uncertainties. That is, $X_z$ and $Y$
induce together a joint probability distribution $p(X,Y|z)$, where the
conditional on $z$ denotes $X_z$. For training, let further assume a finite set $\{
x_i, y_i, z_i \}_{i=1}^N$ of realizations $X_{z_i}(\omega_i), Y(\omega_i)$, for
$\omega_i \in \Omega$ and known values $z_i$ of the nuisance parameter. Our goal
is to learn a function $f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal Y}$ of
parameters $\theta_f$ (e.g., a neural network-based classifier if ${\cal Y}$ is
a finite set of classes) and minimizing  a loss ${\cal L}_f(\theta_f)$ (e.g.,
the cross-entropy). In addition, we require that $f(X_z ; \theta_f)$ should be
robust to the value $z$ of the nuisance parameter  -- which remains unknown at
test time. More specifically, we aim at building $f$ such that in the ideal case
\begin{equation}\label{eqn:criterion-true}
f(X_{z}(\omega) ; \theta_f) = f(X_{z^\prime}(\omega) ; \theta_f)
\end{equation} for all
samples $\omega \in \Omega$ and all $z, z^\prime$ pairs of values of the
nuisance parameter.

@@

    \node[above] at (12.75,-2) {\small $\theta^1(f(X;\theta_f);\theta_r)$};
    \path[-o, shorten >= 0pt] (11,-2.0) edge (14,-2.0);
    \node[above] at (12.75,-3) {$\theta^2(f(X;\theta_f);\theta_r)$};
    \path[-o, shorten >= 0pt] (11,-3) edge (14,-3);
    \node[above] at (12.75,-4) {$\dots$};
    \path[-o, shorten >= 0pt] (11,-4) edge (14,-4);

@@

    \node[neuron] (r-O1) at (7cm+4*\layersep,-2cm) {};
    \node[neuron] (r-O2) at (7cm+4*\layersep,-3cm) {};
    \node[neuron] (r-O3) at (7cm+4*\layersep,-4cm) {};

@@

    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O1);
    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O2);
    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O3);

@@

    \draw (14,-1.5) rectangle (17,-4.5);
    \path[->, shorten >= 0pt] (15.5,-0.5) edge (15.5,-1.5);
    \node[above] at (15.5,-0.5) {$Z$};
    \path[->, shorten >= 0pt] (15.5,-4.5) edge (15.5,-5.5);
    \node[below] at (15.5,-5.5) {$p_{\theta_r}(Z|f(X;\theta_f))$};
    \node at (15.5,-3) {${\cal D}(\theta^1, \theta^2, \dots)$};

@@

    \caption{Architecture for the adversarial training of a binary classifier
    $f$ against a continuous parameter $Z$.}
    \label{fig:architecture}
    \end{figure*}
